---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 8,
  fig.height = 4,
  dpi = 150
)
has_meta <- requireNamespace("meta", quietly = TRUE)
```

# ggmeta

<!-- badges: start -->
[![R-CMD-check](https://github.com/drhrf/ggmeta/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/drhrf/ggmeta/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/drhrf/ggmeta/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/drhrf/ggmeta/actions/workflows/pkgdown.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

**ggmeta** builds publication-quality forest and funnel plots with
[ggplot2](https://ggplot2.tidyverse.org). Give it a `meta` object (from the
[meta](https://cran.r-project.org/package=meta) package) or a plain tidy data
frame — the result is an ordinary `ggplot` you can theme, compose, and save.

## Installation

```r
# install.packages("remotes")
remotes::install_github("drhrf/ggmeta")
```

## Quick start

No `meta` package required — a tidy data frame of effect sizes and standard
errors is enough. Set `add_summary = TRUE` to pool the studies on the fly
(inverse-variance common effect and DerSimonian–Laird random effects):

```{r quick-start}
library(ggmeta)

studies <- data.frame(
  studlab  = c("Trial 1", "Trial 2", "Trial 3", "Trial 4", "Trial 5"),
  estimate = c(0.10, 0.35, 0.22, 0.48, 0.05),
  se       = c(0.12, 0.10, 0.14, 0.16, 0.11)
)
studies$ci_lower <- studies$estimate - 1.96 * studies$se
studies$ci_upper <- studies$estimate + 1.96 * studies$se

ggforest(studies, add_summary = TRUE)
```

## Estimates, CIs, weights, and heterogeneity

Pass a `meta` object and add `columns = TRUE` to reproduce the familiar
`meta::forest()` table: an effect estimate, 95% CI, and weight column for every
study and summary, headers, and a heterogeneity line (*I²*, the between-study
variance, *Q*, and *p*) — all as a plain `ggplot`.

```{r meta-columns, eval = has_meta, fig.width = 9.5, fig.height = 4.2}
library(meta)

dat <- data.frame(
  study   = c("Adams 2019", "Baker 2020", "Chen 2020",
              "Diaz 2021", "Evans 2022", "Foster 2023"),
  event.e = c(12,  8, 25, 18, 30, 15), n.e = c(120,  90, 200, 150, 250, 130),
  event.c = c(20, 14, 30, 28, 35, 25), n.c = c(118,  92, 205, 148, 245, 128)
)

m <- metabin(event.e, n.e, event.c, n.c,
             data = dat, studlab = study, sm = "RR")

ggforest(m, columns = TRUE)
```

Everything is optional. Choose which columns to show, and toggle the other
elements on or off:

```r
ggforest(m, columns = c("estimate", "ci")) # only some columns
ggforest(m, effect_header = "Risk ratio")  # rename the estimate column
ggforest(m, show_hetstats = FALSE)          # hide the heterogeneity line
ggforest(m, show_predict  = FALSE)          # hide the prediction interval
ggforest(m, sort_studies  = FALSE)          # keep the input order
```

## Forest *and* funnel plots on one canvas

`ggmeta` also draws funnel plots with `ggfunnel()` (study effect vs. standard
error, with pseudo confidence-interval contours). And because every plot is an
ordinary `ggplot`, a forest and a funnel compose on a single figure with
[patchwork](https://patchwork.data-imaginist.com) — something that is awkward
with the base-graphics output of `meta`:

```{r patchwork, eval = has_meta, fig.width = 11, fig.height = 4.6}
library(patchwork)

(ggforest(m) | ggfunnel(m)) +
  plot_layout(widths = c(2, 1)) +
  plot_annotation(tag_levels = "A")
```

## Journal styles

Layout presets restyle a plot for common journals. They are ordinary ggplot2
components, so you add them with `+`:

```{r layout-jama, eval = has_meta, fig.width = 9.5, fig.height = 4.2}
layout_jama(ggforest(m, columns = TRUE))
```

`layout_bmj()` and `layout_revman5()` are also available.

## Custom columns

For a column of your own — sample sizes, events, anything — use
`geom_forest_text()`, aligned to the study rows through the shared `y`.
`tidy_meta()` exposes the same tidy data frame `ggforest()` builds internally,
and `format_effect()` builds `"estimate (low to high)"` labels:

```r
td      <- tidy_meta(m)
studies <- td[!td$is_summary, ]

ggforest(m) +
  geom_forest_text(aes(y = studlab, label = n.e), data = studies,
                   x = 4, hjust = 0) +
  expand_limits(x = 6)
```

## Why ggmeta?

- **ggplot2 native** — add themes, layers, annotations, and facets; save with `ggsave()`.
- **Forest and funnel plots** — `ggforest()` and `ggfunnel()`, both ordinary ggplots, so you can arrange them together with patchwork.
- **`meta::forest()`-style tables** — estimate / CI / weight columns and a heterogeneity caption, via `columns = TRUE`.
- **Standalone or `meta`** — works on a tidy data frame or a `meta` object, and can pool studies itself (`add_summary = TRUE`).
- **Correct by construction** — every summary measure is back-transformed with its right inverse (ratios, logit proportions, Fisher-*z* correlations, rates).
- **Custom geometries** — proper `ggproto` geoms for CIs, summary diamonds, prediction intervals, reference lines, and text columns.
- **Journal styles** — `layout_jama()`, `layout_bmj()`, `layout_revman5()`.

## Learn more

- `vignette("getting-started")` — a tour of the package.
- `vignette("customising")` — restyle every element of a forest or funnel plot.
- `vignette("from-meta-forest")` — coming from `meta::forest()`.
