---
title: "Customising forest and funnel plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customising forest and funnel plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 3.6,
  dpi = 96
)
has_meta <- requireNamespace("meta", quietly = TRUE)
```

`ggforest()` and `ggfunnel()` assemble several layers for you 〔 study
confidence intervals, summary diamonds, a prediction interval, reference lines,
funnel points and contours. You restyle any of them with a matching `*_args`
argument: a named list of arguments passed straight to the underlying geom.

This is the intended way to change those elements. Adding another
`geom_forest_*()` layer to a `ggforest()` plot does **not** restyle the built-in
one 〔 it draws a *second* layer over every row.

```{r setup, message = FALSE}
library(ggmeta)
library(ggplot2)
```

```{r model, eval = has_meta}
library(meta)

dat <- data.frame(
  study   = c("Adams 2019", "Baker 2020", "Chen 2020",
              "Diaz 2021", "Evans 2022", "Foster 2023"),
  event.e = c(12,  8, 25, 18, 30, 15), n.e = c(120,  90, 200, 150, 250, 130),
  event.c = c(20, 14, 30, 28, 35, 25), n.c = c(118,  92, 205, 148, 245, 128)
)
m <- metabin(event.e, n.e, event.c, n.c,
             data = dat, studlab = study, sm = "RR")
```

## The prediction interval

`predict_args` controls the prediction interval 〔 `colour`, `linetype`,
`linewidth`, `alpha`, and the end-cap size `cap_width`:

```{r predict, eval = has_meta}
ggforest(m, predict_args = list(
  cap_width = 0.1, colour = "firebrick", linewidth = 0.8, linetype = "solid"
))
```

## Summary diamonds

Recolour the diamonds with `diamond_colours` 〔 a named vector keyed by
`"common"`, `"random"`, `"subgroup_common"`, `"subgroup_random"` 〔 and restyle
their border or transparency with `diamond_args`:

```{r diamonds, eval = has_meta}
ggforest(m,
  diamond_colours = c(common = "grey45", random = "#1B7837"),
  diamond_args    = list(colour = "grey20", alpha = 1)
)
```

## Study intervals and reference lines

`ci_args` styles the study confidence intervals and their weight-proportional
squares (including `point_size_range`). `ref_args` styles the null-effect line,
and `consensus` / `consensus_args` control the dotted pooled-estimate line:

```{r ci-ref, eval = has_meta}
ggforest(m,
  ci_args   = list(colour = "grey30", point_size_range = c(1, 5)),
  ref_args  = list(linetype = "dashed"),
  consensus = FALSE
)
```

## Everything together

The styling arguments combine freely, and work with the `meta::forest()`-style
table columns too:

```{r all, eval = has_meta, fig.width = 9}
ggforest(m, columns = TRUE,
  predict_args    = list(cap_width = 0.1, colour = "firebrick"),
  diamond_colours = c(common = "grey45", random = "#1B7837"),
  ci_args         = list(colour = "grey30")
)
```

## Funnel plots

`ggfunnel()` follows the same pattern with `point_args` (the study points),
`contour_args` (the pseudo confidence-interval contours), and `ref_args` (the
vertical reference line):

```{r funnel, eval = has_meta, fig.width = 6, fig.height = 4.6}
ggfunnel(m,
  point_args   = list(size = 3, fill = "#1B7837"),
  contour_args = list(colour = "grey70", linetype = "dotted", level = c(0.95, 0.99)),
  ref_args     = list(colour = "firebrick")
)
```

## See also

- `vignette("getting-started")` 〔 a tour of the package.
- `vignette("from-meta-forest")` 〔 coming from `meta::forest()`.
- `?ggforest` and `?ggfunnel` 〔 the full list of styling arguments.
